### 2018-01-26 Draft/abstract.tex
series of imporvements, corrections from the beginning up to section 3.3

\begin{abstract}
\noindent

In this article, a detailed study of the vector-boson scattering for two positively charged W bosons is presented.
In particular, a comparison between the full next-to-leading (NLO) QCD corrections against several approximations is carried out.
This study is not only performed in the usual fiducial region used by experimental collaborations but also in a more inclusive set-up.
This allows to infer precisely the quality of such approximations.
Finally, NLO predictions matched to various parton shower are also discussed.
Thanks to this, it is thus possible to infer the systematic errors related to vector-boson scattering at the NLO-QCD level and beyond.

\end{abstract}
\thispagestyle{empty}
\vfill
\newpage
\setcounter{page}{1}

%\tableofcontents
\newpage