### 2018-02-08 Draft/abstract.tex
Add my corrections

\begin{abstract}
  \noindent

  Vector-boson scattering processes are of great importance for the
  current run-2 and future runs at design energy of the Large Hadron
  Collider. The presence of triple and quartic gauge couplings gives
  access to tests of the gauge sector and possible new-physics
  contributions there. For this, a precise knowledge of the Standard
  Model contributions is necessary, which at least matches the
  experimental uncertainties.

  In this article, we present a detailed study of the vector-boson
  scattering process with two positively charged leptons and missing
  transverse momentum in the final state, mediated predominantly by
  same-sign production of two W bosons with positive charge.
In particular, we carry out a comparison between the full NLO QCD corrections against several approximations.
This study is not only performed in the usual fiducial region used by experimental collaborations, but also in a more inclusive set-up.
This allows to infer precisely the quality of such approximations.
Finally, we discuss NLO predictions matched to various parton showers.
Thanks to this, it is thus possible to infer the systematic errors related to vector-boson scattering at the NLO-QCD level and beyond.

\end{abstract}
\thispagestyle{empty}
\vfill
\newpage
\setcounter{page}{1}

%\tableofcontents
\newpage