### 2018-02-15 Draft/abstract.tex
direct modifications in the gitlab, please check!

\begin{abstract}
  \noindent

  Vector-boson scattering processes are of great importance for the
  current run-2 and future runs at design energy of the Large Hadron
  Collider. The presence of triple and quartic gauge couplings in the process
  gives access the gauge sector of the Standard Model and possible new-physics
  contributions there. 
  To test this hypothesis, a precise knowledge of the Standard
  Model contributions is necessary, 
  with a precision which at least matches the
  experimental uncertainties
  of existing and forthcoming measurements.

  In this article, we present a detailed study of the vector-boson
  scattering process with two positively charged leptons and missing
  transverse momentum in the final state, mediated predominantly by
  same-sign production of two W bosons with positive charge.
In particular, 
we carry out a comparison between the full fixed-order NLO QCD corrections 
against several approximations.
This study is performed in the usual fiducial region used by experimental collaborations, 
as well as in a more inclusive phase space,
where the importance of the approximations can be infered.
Eventually, we discuss NLO predictions matched to various parton showers.
In summary, from the distributions presented it is thus possible to infer 
the systematic unertainties due to missing higher orders in the calculations
at the NLO + NLL QCD level.

\end{abstract}
\thispagestyle{empty}
\vfill
\newpage
\setcounter{page}{1}

%\tableofcontents
\newpage