### 2018-02-19 Draft/acknowledgements.tex
added acknowledgments

We would like to thank \ldots \\
the {\sc Pythia8} authors, in particular Stefan Prestel, Torbjorn Sjostrand and Peter Skands for 
discussions and clarifications about the third-jet rapidity spectrum.

The authors would like to acknowledge the contribution of the COST Action CA16108 which initiated this work.
Moreover, this work was supported by several STSM Grant from the COST Action CA16108. Many authors acknowledge
hospitality from Nikhef, where part of this work was performed.

BB, AD, and MP acknowledge financial support by the
German Federal Ministry for Education and Research (BMBF) under
contract no.~05H15WWCA1 and the German Science Foundation (DFG) under
reference number DE 623/6-1.

MZ is supported
 by the Netherlands National Organisation for Scientific
 Research (NWO),