### 2018-02-20 Draft/scanLO.tex

In Fig.~\ref{fig:ratio2d_LO}, ratios of double-differential cross sections in the plane $\left(m_{\Pj\Pj}, \Delta y_{\Pj\Pj}\right)$ is shown.
Two plots are displayed: the ratios of the $|t|^2 + |u|^2$ and $|s|^2 + |t|^2 + |u|^2$ approximations over the full calculation.
In the first case, the approximation is good within $\pm10\%$ over the whole range apart in the low invariant-mass region at both low and large rapidity difference.
The low rapidity difference region possesses remnants of the tri-bosons contributions that peak at di-jet invariant mass around the W-boson mass.
It is therefore expected that the $|t|^2 + |u|^2$ approximation fails in this region.
The second plot, where the $|s|^2 + |t|^2 + |u|^2$ approximation is considered, displays a better behaviour in the previously mentioned region.
The full calculation is approximated at the level of $\pm5\%$ apart in the region where $\Delta y_{\Pj\Pj} < 2$.
\AK{What are the MC error on these plots?}
\MP{@AK: I changed some stuffs in the text to justify the plot range}

\begin{figure*}[hbt]
\centering
\includegraphics[scale=0.395]{figures/scanfigures/ratio_tu.pdf}
\includegraphics[scale=0.395]{figures/scanfigures/ratio_stu.pdf}
\caption{Ratio of cross sections per bin in the plan $\left(m_{\Pj\Pj}, |\Delta y_{\Pj\Pj}| \right)$ at LO \emph{i.e.}\ order $\mathcal{O}(\alpha^6)$.
Ratio of approximated squared amplitudes over the full matrix element.
The approximated squared amplitudes are computed as $|\mathcal{A}|^2 \sim |t|^2 + |u|^2$ (left) and $|\mathcal{A}|^2 \sim |s|^2 + |t|^2 + |u|^2$ (right).
The cuts applied are the one of of Sec.~\ref{subsec:inputpar} and no cuts on $m_{\Pj\Pj}$ and $|\Delta y_{\Pj\Pj}|$ are applied.} 
\label{fig:ratio2d_LO}
\end{figure*}
% As explained previously, the low di-jet invariant mass and low jet rapidity separation regions are dominated by tri-boson production.

Therefore, the \emph{inclusive} study at NLO is only performed in the region 
%
\begin{equation}
\label{eq:inclusive}
	m_{jj} > 200\GeV\, \qquad {\rm and} \qquad |\Delta y_{jj}| > 2 .
\end{equation}
%
Hence, the differences arising at NLO in this fiducial region originate solely from NLO effects.

% As explained previously, the low di-jet invariant mass are dominated by tri-boson production.
% Therefore, a comparison of the various approximations in an \emph{inclusive} phase-space volume should exclude the region where tri-boson contributions are dominating.
% To that end, we have chosen for the inclusive fiducial volume to take the following values for the VBS cuts:
%