### 2018-02-20 Draft/crosssectionLO.tex

In Tab.~\ref{tab:wg1_LOrates}, we report the total rates at LO accuracy obtained with the set-up described in Eqs.~(\ref{cut:1}-\ref{cut:3}) with the VBS cuts $m_{\Pj\Pj} > 500 \GeV$ and $|\Delta y_{\Pj\Pj}| > 2.5$ (see Eq.~(\ref{cut:4})).
The order considered here is the order $\mathcal{O}(\alpha^6)$.
We note that several full predictions are not in statistical agreement. These are possibly due to Monte-Carlo 
integrators performing too aggressive estimations of statistical uncertainties. Nonetheless, all these predictions agree within less than $0.5\%$.
At the level of the cross section, it seems difficult to infer the quality of the the various approximations.
This simply means that the details of the various VBS approximations have an impact not larger than $0.5\%$ at 
the level of the fiducial cross section at LO for a typical phase-space volume used by experimental collaborations. This is in agreement with the findings of Ref.~\cite{Denner:2012dz,Oleari:2003tc}.

\begin{table}[h!]
    \centering
    \begin{tabular}{c|r@{ $\pm$ }l}
      Code  &  \multicolumn{2}{c}{$\sigma[\rm{fb}]$}  \\
        \hline
        \hline
        {\sc Bonsay}  &  $1.43636$ & $0.00002$ \\
        {\sc MG5\_aMC}&  $1.430\phantom{0}\phantom{0}$ & $0.001$ \\
        {\sc MoCaNLO+Recola}  &  $1.43476$ & $0.00009$ \\
        {\sc PHANTOM} &  $1.4374\phantom{0}$ & $0.0006 $  \\
        {\sc Powheg-Box}  &  $1.44092$ & $0.00009$ \\
        {\sc VBFNLO}  &  $1.43796$ & $0.00005$ \\
        {\sc Whizard} &  $1.4381\phantom{0}$ & $0.0002 $
    \end{tabular}
    \caption{\label{tab:wg1_LOrates} Cross sections for the LHC process ${\rm p}{\rm p}\to\mu^+\nu_\mu{\rm e}^+\nu_{\rm e}{\rm j}{\rm j}$ at LO accuracy and order $\mathcal{O}(\alpha^6)$.
    The uncertainties shown refer to the estimated statistical error of the Monte Carlo programs.
    The predictions are obtained in the fiducial region described in Sec.~\ref{subsec:inputpar}.
    \MP{Please add or check your respective numbers}}
\end{table}