### 2018-03-04 Draft/acknowledgements.tex
Significant modifications of Mathieu: corrections/improvements, all labels should now be consistent

The authors thank the {\sc Pythia8} authors, in particular Stefan Prestel, Torbjorn Sjostrand and Peter Skands for 
discussions and clarifications about the third-jet rapidity spectrum.

The authors would like to acknowledge the contribution of the COST Action CA16108 which initiated this work.
Moreover, this work was supported by several STSM Grants from the COST Action CA16108.
Many authors acknowledge hospitality from Nikhef, where part of this work has been performed.

BB, AD, and MP acknowledge financial support by the
German Federal Ministry for Education and Research (BMBF) under
contract no.~05H15WWCA1 and the German Science Foundation (DFG) under
reference number DE 623/6-1.
AK acknowledges financial support by the Swiss National Science Foundation (SNF) under contract 200020-175595.
MR acknowledges funding from the European Union's Horizon 2020 research and innovation programme as part of the Marie Sklodowska-Curie Innovative Training Network MCnetITN3 (grant agreement no. 722104).
CS acknowledges support by the state of Baden-W\"urttemberg through bwHPC and the German Research Foundation (DFG) through grant no INST 39/963-1 FUGG and grant DI 784/3.
HSS is supported by the ILP Labex (ANR-11-IDEX-0004-02, ANR-10-LABX-63).
MZ is supported by the Netherlands National Organisation for Scientific Research (NWO). The work of BJ was supported in part by the Institutional Strategy of the University of T\"ubingen (DFG, ZUK 63) and in part by
the BMBF under contract number 05H2015.