### 2018-03-04 Draft/crosssectionLO.tex

In Tab.~\ref{tab:wg1_LOrates}, we report the total rates at LO accuracy at order $\mathcal{O}(\alpha^6)$ obtained in the fiducial region described in Sec.~\ref{subsec:inputpar}.
We note that several full predictions are not in statistical agreement. These are possibly due to Monte-Carlo 
integrators performing too aggressive estimations of statistical uncertainties. Nonetheless, all these predictions agree within less than $0.5\%$.
At the level of the cross section, it seems difficult to infer the quality of the the various approximations.
This simply means that the details of the various VBS approximations have an impact not larger than $0.5\%$ at 
the level of the fiducial cross section at LO for a typical phase-space volume used by experimental collaborations.
This is in agreement with the findings of Ref.~\cite{Denner:2012dz,Oleari:2003tc}.

\begin{table}[h!]
    \centering
    \begin{tabular}{c|r@{ $\pm$ }l}
      Code  &  \multicolumn{2}{c}{$\sigma[\rm{fb}]$}  \\
        \hline
        \hline
        {\sc Bonsay}  &  $1.43636$ & $0.00002$ \\
        {\sc MG5\_aMC}&  $1.4304\phantom{0}$ & $0.0007$ \\
        {\sc MoCaNLO+Recola}  &  $1.43476$ & $0.00009$ \\
        {\sc PHANTOM} &  $1.4374\phantom{0}$ & $0.0006 $  \\
        {\sc Powheg-Box}  &  $1.44092$ & $0.00009$ \\
        {\sc VBFNLO}  &  $1.43796$ & $0.00005$ \\
        {\sc Whizard} &  $1.4381\phantom{0}$ & $0.0002 $
    \end{tabular}
    \caption{\label{tab:wg1_LOrates} Cross sections at LO accuracy and order $\mathcal{O}(\alpha^6)$.
    The predictions are obtained in the fiducial region described in Sec.~\ref{subsec:inputpar}.
    The uncertainties shown refer to the estimated statistical error of the Monte Carlo programs.}
\end{table}