### 2018-03-08 Draft/scanNLO.tex
Rephrasing in section 5.1

According to the results of Secs.~\ref{subsec:contributions} and \ref{subsec:LOinclusive}, the VBS approximation at LO fails drastically in the region $m_{\Pj\Pj} < 200$ GeV, $|\Delta y_{\Pj\Pj}| < 2$.
Therefore, we present an inclusive study at NLO QCD for the EW component, namely the order $\mathcal{O}(\alphas\alpha^6)$ for the set-up described in Sec.~\ref{subsec:inputpar} but imposing the requirements $m_{\Pj\Pj}>200 \GeV$ and $|\Delta y_{\Pj\Pj}|>2$.

We compare three different predictions at NLO QCD: 
the VBS approximation implemented in {\sc Bonsay} (dubbed $|t|^2+|u|^2$), the VBS approximation with the $s$-channel contributions from {\sc VBFNLO} (dubbed $|s|^2+|t|^2+|u|^2$), and the full computation.
The full computation employs exact matrix elements meaning that $t/u/s$ interferences, factorisable and non-factorisable QCD corrections, as well as EW corrections to the order $\mathcal{O}(\alphas \alpha^5)$ are included.

The total cross sections within the above mentioned kinematic cuts are shown in Tab.~\ref{tab:crosssecINCLUSIVE}.
The $|t|^2+|u|^2$ approximation for NLO QCD predictions is lower by about $6\%$ than the full calculation.
The inclusion of $s$-channel diagrams improves the approximate prediction, leading to an excess at the $3\%$-level.

\begin{table}[h!]
\centering
\begin{tabular}{c|c|c}
Prediction & $\sigma_{\textrm{tot}}\,[\textrm{fb}]$ & $\delta [\%]$ \\
\hline
\hline
full &  $1.733\phantom{0} \pm 0.002\phantom{0}$ & - \\
\hline
$|t|^2 + |u|^2$ & $1.6292 \pm 0.0001$  &  $-6.0$ \\
\hline
$|s|^2 + |t|^2 + |u|^2$ & $1.7780 \pm 0.0001$  & $+2.6$
\end{tabular}
\caption{Cross sections at NLO QCD \emph{i.e.}\ at order $\mathcal{O}(\alphas\alpha^6)$ for the full computation and two approximations.
In addition to the cuts of Sec.~\ref{subsec:inputpar}, the VBS cuts take the values $m_{\Pj\Pj}>200 \GeV$ and $|\Delta y_{\Pj\Pj}|>2$.
The uncertainties shown refer to the estimated statistical error of the Monte Carlo programs.}
\label{tab:crosssecINCLUSIVE}
\end{table}

These differences are more evident in differential distributions.
In Fig.~\ref{fig:mjjdyjj_1d_1}, we show the differential distributions in the di-jet invariant mass $m_{\Pj\Pj}$ and rapidity separation $|\Delta y_{\Pj\Pj}|$.
For large $m_{\Pj\Pj}$ and large $|\Delta y_{\Pj\Pj}|$, as expected, the VBS approximation is performing well and its $s$-channel extension agrees with the full calculation within $10\%$ per cent.
This is in contrast with the region  $200 \GeV < m_{\Pj\Pj} < 500 \GeV$ and $2<|\Delta y_{\Pj\Pj}|<2.5$, where
the difference between the $|t|^2+|u|^2$ approximation and the full computation can be above $30\%$.
The inclusion of $s$-channels cures partly this behaviour by rendering the approximation good by about $10\%$.
This tends to indicate that interference contributions and/or non-factorisable QCD corrections play a non-negligible role in this phase-space region.

\begin{figure*}[hbt]
\centering
{\includegraphics[scale=0.35]{figures/scanfigures/mjj_nlo.pdf}}
{\includegraphics[scale=0.35]{figures/scanfigures/dyjj_nlo.pdf}}
\caption{
Differential distributions in the di-jet invariant mass (left) and the rapidity-separation of the two tagging jets (right) at NLO QCD \emph{i.e.}\ at order $\mathcal{O}(\alphas\alpha^6)$ for the full computation and two approximations.
The upper plots provide the absolute value for each prediction while the lower plots presents all predictions normalised to {\sc MoCaNLO}+{\sc Recola} which is one of the full predictions.
In addition to the cuts of Sec.~\ref{subsec:inputpar}, the VBS cuts take the values $m_{\Pj\Pj}>200 \GeV$ and $|\Delta y_{\Pj\Pj}|>2$.} 
\label{fig:mjjdyjj_1d_1}
\end{figure*}

In order to investigate further the jet-pair kinematics, we study the double-differential distribution in the variables $m_{\Pj\Pj}$ and $|\Delta y_{\Pj\Pj}|$.
In particular, in Fig.~\ref{fig:ratio2d_NLO}, we compute in each bin the ratios of the approximated cross sections over the full one [$\sigma(|t|^2+|u|^2)/\sigma(\textrm{full})$ and $\sigma(|s|^2+|t|^2+|u|^2)/\sigma(\textrm{full})$].
As expected, in the low invariant-mass and low rapidity-separation region of the jet pair ($200 \GeV < m_{\Pj\Pj} < 500 \GeV$, $2<|\Delta y_{\Pj\Pj}|<2.5$) the VBS approximation fails significantly (by more than $40\%$).
Including the $s$-channel contributions leads to a difference of less than $10\%$ in this very region.
However, in the region of large di-jet invariant mass and low rapidity separation of the jets, the $|s|^2+|t|^2+|u|^2$ approximation overestimate the full computation by more than $40\%$.\footnote{The bin
in the top-left corner of the right-hand side plot of Fig.~\ref{fig:ratio2d_NLO} suffers from large uncertainty ($30\%$) while the other errors are at the per-cent level.}
Again, this seems to support the fact that interferences and non-factorisable corrections can be non-negligible in this region.
On the other hand, in the typical VBS region, the VBS approximation shows a good agreement with the full computation as documented in details in Sec.~\ref{sec:fidNLO}.

\begin{figure*}[hbt]
\centering
{\includegraphics[scale=0.395]{figures/scanfigures/a6as_vbfnloVSrecola_tu.pdf}}
{\includegraphics[scale=0.395]{figures/scanfigures/a6as_vbfnloVSrecola_stu.pdf}}
\caption{Ratios of double-differential distributions in the variables $m_{\Pj\Pj}$ and $|\Delta y_{\Pj\Pj}|$ at NLO QCD \emph{i.e.}\ at order $\mathcal{O}(\alphas\alpha^6)$ for the VBS approximation over the full computation.
Ratio of approximated squared amplitudes over the full matrix element.
The approximated squared amplitudes are computed as $|\mathcal{A}|^2 \sim |t|^2+|u|^2$ (left) and $|\mathcal{A}|^2 \sim |s|^2+|t|^2+|u|^2$ (right)
In addition to the cuts of Sec.~\ref{subsec:inputpar}, the VBS cuts take the values $m_{\Pj\Pj}>200 \GeV$ and $|\Delta y_{\Pj\Pj}|>2$.}
\label{fig:ratio2d_NLO}
\end{figure*}

In Fig.~\ref{fig:mjjdyjj_1d_2}, the distributions in the transverse momentum of the hardest jet and its rapidity are shown.
At low transverse momentum, $|t|^2+|u|^2$ and $|s|^2+|t|^2+|u|^2$ approximations are lower and higher than the full computation by about $20\%$, respectively.
At high transverse momentum, they have a similar behaviour.
They both diverge from the full computation towards larger transverse momentum to reach about $10\%$ at $1000\GeV$.
Regarding the rapidity of the hardest jet, the two approximations have opposite behaviours.
In the central region, the $|t|^2+|u|^2$ approximation differs by $12\%$ with respect to the full computation while the $|s|^2+|t|^2+|u|^2$ one is good within $5\%$.
In the peripheral region, the $|t|^2+|u|^2$ approximation is rather close to the full computation ($5\%$) while the $|s|^2+|t|^2+|u|^2$ one differs by $10\%$.

\begin{figure*}[hbt]
\centering
{\includegraphics[scale=0.35]{figures/scanfigures/ptj1_nlo.pdf}}
{\includegraphics[scale=0.35]{figures/scanfigures/yj1_nlo.pdf}}
\caption{Differential distributions in the transverse momentum (left) and rapidity of the hardest tagging jet (right) at NLO QCD \emph{i.e.}\ at order $\mathcal{O}(\alphas\alpha^6)$ for the full computation and two approximations.
The upper plots provide the absolute value for each prediction while the lower plots presents all predictions normalised to {\sc MoCaNLO}+{\sc Recola} which is one of the full predictions.
In addition to the cuts of Sec.~\ref{subsec:inputpar}, the VBS cuts take the values $m_{\Pj\Pj}>200 \GeV$ and $|\Delta y_{\Pj\Pj}|>2$.} 
\label{fig:mjjdyjj_1d_2}
\end{figure*}

Concerning leptonic observables, we show in Fig.~\ref{fig:mjjdyjj_1d_3} the distributions in the di-lepton invariant mass and in the Zeppenfeld variable of the electron, defined as
%
\begin{equation}
  z_{\Pe^+} = \frac{y_{\Pe^+}-\frac{y_{\Pj_1}+y_{\Pj2}}2}{|\Delta y_{jj}|} \,.
  \label{eq:Zeppenfeld}
\end{equation}
%
Analogous definitions will later also be used for the Zeppenfeld variable of the muon and of the third jet.
%
\begin{figure*}[hbt]
\centering
{\includegraphics[scale=0.35]{figures/scanfigures/mll_nlo.pdf}}
{\includegraphics[scale=0.35]{figures/scanfigures/zel_nlo.pdf}}
\caption{Differential distributions in the lepton-lepton invariant mass (left) and the electron Zeppenfeld variable (right) at NLO QCD \emph{i.e.}\ at order $\mathcal{O}(\alphas\alpha^6)$ for the full computation and two approximations.
The upper plots provide the absolute value for each prediction while the lower plots presents all predictions normalised to {\sc MoCaNLO}+{\sc Recola} which is one of the full predictions.
In addition to the cuts of Sec.~\ref{subsec:inputpar}, the VBS cuts take the values $m_{\Pj\Pj}>200 \GeV$ and $|\Delta y_{\Pj\Pj}|>2$.} 
\label{fig:mjjdyjj_1d_3}
\end{figure*}
%
The $|s|^2+|t|^2+|u|^2$ predictions for $m_{\rm e^+\mu^+}$ agrees rather well with the full curve, obtained from {\sc MoCa\-NLO+Recola}.
The prediction from {\sc Bonsay} is about $10\%$ lower around $1000 \GeV$.
The Zeppenfeld variable of the positron $z_e$ is more strongly affected by the exclusion of $s$-channel contributions.
For increasing $z_{\rm e}$, the $|t|^2+|u|^2$ approximation diverges from the full computation to reach a difference of about $25\%$ at $1.5$.
On the other hand, including $s$-channel contributions leads to a better approximation, staying within $10\%$ difference over the whole range.

In conclusion, both the loose minimum di-jet invariant mass cut and the inclusion of QCD radiative correction make the $s$-channel contributions less suppressed than at LO, making their inclusion mandatory, in order to provide trustworthy predictions at NLO accuracy.
In the inclusive region studied here, neglecting $s$-channel, non-factorisable corrections, and electroweak corrections can lead to discrepancies of up to $30\%$ with respect to the full computation.
Nevertheless, the VBS approximation at NLO provides a good approximation of full calculations in the kinematic region where VBS contributions are dominant ($m_{\Pj\Pj} \gtrsim 500 \GeV$, $|\Delta y_{\Pj\Pj}| \gtrsim 2.5$), for both total cross section and differential distributions.
This more exclusive region is studied in more details in the next section.