### 2018-03-12 Draft/article.tex
Clean-up of the .tex files

%% Main-Version Draft
\documentclass[twocolumn,epjc3]{svjour3} % use the EPJC template, two columns
%%\documentclass[epjc3]{svjour3} % use the EPJC template, one column
\smartqed  % flush right qed marks, e.g. at end of proof
%% -------------------------------
%% |          Packages           |
%% -------------------------------
 \usepackage{amsmath}
 \usepackage{amssymb}
 \usepackage{graphicx}
  \usepackage[merge,numbers,compress]{natbib}
 \usepackage[T1]{fontenc}
 \usepackage{booktabs}
 \usepackage{xcolor}
 \usepackage{xspace}
 \usepackage{dcolumn}
 \usepackage{hyperref}
%   \usepackage[draft]{hyperref}
 \usepackage{caption}
 \usepackage{tabularx}
 \usepackage{lmodern}
 \usepackage{url}
 \usepackage{morefloats}
 \usepackage[utf8]{inputenc}
%  \usepackage{siunitx}
% \usepackage
% [subrefformat=parens,position=top,skip=-15pt,margin=15pt,justification=justified,singlelinecheck=false]
% {subcaption}
\hyphenation{quan-ti-ta-tive-ly fer-mi-on fer-mi-ons}

%\setlength{\evensidemargin}{0cm}
%\setlength{\oddsidemargin}{0cm}
%\setlength{\topmargin}{0.00cm}
%\setlength{\textwidth}{16.0cm}
%\setlength{\textheight}{22.55cm}
%\setlength{\headheight}{0cm}
%\setlength{\headsep}{0cm}
%\setlength{\voffset}{0cm}
%\setlength{\paperheight}{27cm}

%\renewcommand{\topfraction}{0.8}
%\renewcommand{\bottomfraction}{0.5}
%\renewcommand{\textfraction}{0.2}
%\renewcommand{\floatpagefraction}{0.7}

\newcommand{\MP}[1]{{ {\color{blue}{ [MP: #1]}} }}
\newcommand{\MR}[1]{{ {\color{red}{ [MR: #1]}} }}
\newcommand{\AK}[1]{{ {\color{magenta}{ [AK: #1]}} }}
\newcommand{\ZM}[1]{{ {\color{green}{ [MZ: #1]}} }}
\newcommand{\BJ}[1]{{ {\color{cyan}{ [BJ: #1]}} }}
\newcommand{\GP}[1]{{ {\color{orange}{ [GP: #1]}} }}

%% ---------------------------------
%% | ToDo Marker - only for draft! |
%% ---------------------------------
% Remove this section for final version!
%\setlength{\marginparwidth}{20mm}

\newcommand{\margtodo}
{\marginpar{\textbf{\textcolor{red}{ToDo}}}{}}

\newcommand{\todo}[1]
{{\textbf{\textcolor{red}{(\margtodo{}#1)}}}{}}

\input{macros}
%
\journalname{DESY 18-025, FR-PHENO-2018-003, KA-TP-05-2018, UWTHPH-2018-12, ZU-TH-08/18} % Preprint numbers here
%
\begin{document}

\title{Precise predictions for same-sign W-boson scattering at the LHC}

\author{Alessandro Ballestrero\thanksref{infnto}
\and
Benedikt Biedermann\thanksref{wurz}
\and
Simon Brass\thanksref{sieg}
\and
Ansgar Denner\thanksref{wurz}
\and
Stefan Dittmaier\thanksref{frei}
\and
Rikkert Frederix\thanksref{tum}
\and
Pietro Govoni\thanksref{mila}
\and
Michele Grossi\thanksref{pavia,ibm}
\and
Barbara J\"ager\thanksref{tub}
\and
Alexander Karlberg\thanksref{uzh}
\and
Ezio Maina\thanksref{infnto,unito}
\and
Mathieu Pellen\thanksref{wurz}
\and
Giovanni Pelliccioli\thanksref{infnto,unito}
\and
Simon Pl\"atzer\thanksref{wien}
\and
Michael Rauch\thanksref{kit}
\and
Daniela Rebuzzi\thanksref{pavia}
\and
J\"urgen Reuter\thanksref{desy}
\and
Vincent Rothe\thanksref{desy}
\and
Christopher Schwan\thanksref{frei}
\and
Hua-Sheng Shao\thanksref{lpthe}
\and
Pascal Stienemeier\thanksref{desy}
\and
Giulia Zanderighi\thanksref{cern}
\and
Marco Zaro\thanksref{nikh}
\and
Dieter Zeppenfeld\thanksref{kit}
}

\institute{INFN, Sezione di Torino, %
Via P. Giuria 1, %
10125 Turin, %
Italy\label{infnto}
\and
Universit\"at W\"urzburg, %
Institut f\"ur Theoretische Physik und Astrophysik,  %
Emil-Hilb-Weg 22, %
97074 W\"urzburg, %
Germany\label{wurz}
\and
Universit\"at Siegen, Department Physik, %
 Walter-Flex-Str.3, %
57068 Siegen, Germany\label{sieg}
\and
Albert-Ludwigs-Universit\"at Freiburg, Physikalisches Institut, %
 Hermann-Herder-Str.\ 3, %
79104 Freiburg, Germany\label{frei}
\and
Technische Universit\"{a}t M\"{u}nchen, %
James-Franck-Str.~1, %
85748 Garching, %
Germany\label{tum}
\and
University and INFN, Milano-Bicocca, %
piazza della Scienza, 3, 20126 Milan, %
Italy\label{mila}
\and
Universit\'a di Pavia, Dipartimento di Fisica and INFN, Sezione di Pavia, %
Via A. Bassi 6, %
27100 Pavia, %
Italy\label{pavia}
\and
IBM Italia s.p.a. %
Circonvallazione Idroscalo, %
20090 Segrate (MI), %
Italy\label{ibm}
\and
University of T\"ubingen,
Institute for Theoretical Physics,
Auf der Morgenstelle 14,
72076 T\"ubingen,
Germany\label{tub}
\and
Universit\"at Z\"urich,
Physik-Institut,
Winterthurerstrasse 190,
8057 Z\"urich,
Switzerland\label{uzh}
\and
Universit\`a di Torino, %
Dipartimento di Fisica, %
Via P. Giuria 1, %
10125 Turin, %
Italy\label{unito}
\and
University of Vienna,
Particle Physics, Faculty of Physics, %
Vienna, %
Austria\label{wien}
\and
Karlsruhe Institute of Technology (KIT), %
Institute for Theoretical Physics, %
76131 Karlsruhe,
Germany\label{kit}
\and
DESY Theory Group, %
Notkestr. 85, %
22607 Hamburg,
Germany\label{desy}
\and
Sorbonne Universit\'e et CNRS, %
Laboratoire de Physique Th\'eorique et Hautes \'Energies (LPTHE), % 
UMR 7589, 
4 place Jussieu, %
75252 Paris Cedex 05, %
France\label{lpthe}
\and
CERN,
Theoretical Physics Department,
1211, Geneva 23,
Switzerland\label{cern}
\and
Nikhef,
Science Park 105,
1098XG Amsterdam,
The Netherlands\label{nikh}
}

\maketitle

\input{abstract}

\section{Introduction}
\input{introduction}

\section{Definition of the process}
    \label{sec:definition}
    \input{definition}

\section{Details of the calculations}
    \label{sec:details}
    \subsection{Several descriptions for one process}
        \input{process}
    \subsection{Description of the predictions}
        \label{subsec:codedescr}
        \input{description}
    \subsection{Input parameters}
        \label{subsec:inputpar}
        \input{input}

\section{Leading-order study}
    \label{sec:LO}
    \subsection{Contributions}
        \label{subsec:contributions}
        \input{contributions}
    \subsection{Inclusive comparison}
        \label{subsec:LOinclusive}
        \input{scanLO}
    \subsection{Comparison in the fiducial region}
        \label{subsec:LOfiducial}
        \input{crosssectionLO}
        \input{distributionsLO}

\section{Next-to-leading order QCD}
    \label{sec:NLO}
    \subsection{Inclusive comparision}
        \label{subsec:NLOinclusive}
        \input{scanNLO}
    \subsection{Comparison in the fiducial region}
    \label{sec:fidNLO}
        \input{crosssectionNLO}
        \input{distributionsNLO}

\section{Matching to parton shower}
    \label{sec:matching}
    \input{shower}

\section{Conclusions and recommendations}
\label{sec:conclusion}

\input{conclusion}

\section*{Acknowledgements}

\input{acknowledgements}

\bibliographystyle{utphys}
\bibliography{article}

\end{document}