### 2018-03-12 Draft/old_tex/abstract.tex

\begin{abstract}
  \noindent

  Vector-boson scattering processes are of great importance for the
  current run-II and future runs of the Large Hadron
  Collider. The presence of triple and quartic gauge couplings in the
  process gives access to the gauge sector of the Standard Model and
  possible new-physics contributions there.  To test this hypothesis,
  sound knowledge of the Standard Model contributions is necessary,
  with a precision which at least matches the experimental
  uncertainties of existing and forthcoming measurements.  In this
  article we present a detailed study of the vector-boson scattering
  process with two positively-charged leptons and missing transverse
  momentum in the final state, mediated predominantly by same-sign
  production of two $\PW$ bosons with positive charge.  In particular,
  we first carry out a systematic comparison of the various
  approximations that are usually performed for this kind of process
  against the complete calculation, at LO and NLO QCD accuracy. Such a
  study is performed both in the usual fiducial region used by
  experimental collaborations and in a more inclusive phase space,
  where the differences among the various approximations lead to more
  sizeable effects.  Afterwards, we turn to predictions
  matched to parton showers, at LO and NLO: we show that on the one
  hand, the inclusion of NLO QCD corrections lead to more stable
  predictions, but on the other the details of the matching and of the
  parton-shower programs lead to differences which are considerably
  larger than those observed at fixed-order, even in the experimental
  fiducial region.  We conclude with some recommendations for
  experimental studies of vector-boson scattering processes.

\end{abstract}
\thispagestyle{empty}
\vfill
\newpage
\setcounter{page}{1}

%\tableofcontents
\newpage